基于卷积神经网络的视频目标跟踪算法研究

### 2021-02-23 博士/毕业论文/ucasthesis/Tex/Frontmatter.tex
润色 intro
更新英文摘要

目标跟踪是计算机视觉领域中最重要和最具挑战性的研究课题之一，在智能监控、自动驾驶等领域有着潜在广阔的应用前景。目标跟踪的核心是估计图像序列的每帧中目标的运动状态。目标跟踪是计算机视觉领域的中层部分，为目标的行为理解提供了基础，因此具有非常重要的理论研究价值。同时，它具有广泛的实际应用，包括视频监控，交通流量监控，视频压缩和人机交互等。
%例如，目标跟踪已成功应用于监控居民区，停车场和银行中的人类活动（例如W4系统[1]和VSAM项目[2]）。在交通运输领域，目标跟踪也被广泛用于交通流量监控[3]，行人计数[4]等任务。由于目标跟踪的理论价值与应用价值，众多科研机构和公司都投入到这项研究中。
然而，目标跟踪领域存在很多理论和技术问题有待解决，如运动模糊、光照变化、非刚性目标的形变、视角的变化导致的目标旋转、遮挡等。近年来深度学习的突破为解决目标跟踪中的一系列问题带来了可能。
深度学习是基于人工神经网络的机器学习方法。在过去的十年中，深度学习技术得到了飞速发展，已成功应用于计算机视觉，语音识别，自然语言处理，音频识别，社交网络过滤，机器翻译，生物信息学，药物设计等领域。如何利用深度学习方法，尤其是深度卷积神经网络解决跟踪过程中遇到的复杂问题，具有较大的研究价值和研究空间。

@@

\keywords{相关滤波，视频目标跟踪，深度学习，卷积神经网络}% 中文关键词
%-
%-> 英文摘要
%-
\intobmk\chapter*{Abstract}% 显示在书签但不显示在目录

Target tracking is one of the most important and challenging research topics in the field of computer vision, and it has potential and broad application prospects in the fields of intelligent surveillance and autonomous driving. The core of target tracking is to estimate the motion state of the target in each frame of the image sequence. Target tracking is the middle part of the computer vision field, which provides a basis for the understanding of target behavior, so it has very important theoretical research value. At the same time, it has a wide range of practical applications, including video surveillance, traffic flow monitoring, video compression, and human-computer interaction. For example, target tracking has been successfully applied to monitor human activities in residential areas, parking lots and banks (such as the W4 system [1] and the VSAM project [2]). In the field of transportation, target tracking is also widely used in traffic flow monitoring [3], pedestrian counting [4] and other tasks.
Due to the theoretical value and application value of target tracking, many scientific research institutions and companies have invested in this research. However, there are many theoretical and technical problems to be solved in the field of target tracking, such as motion blur, illumination changes, deformation of non-rigid targets, target rotation and occlusion caused by changes in viewing angle. In recent years, the breakthrough of deep learning has brought the possibility to solve a series of problems in target tracking.
Deep learning is a machine learning method based on artificial neural networks. In the past ten years, deep learning technology has been rapidly developed and has been successfully applied in computer vision, speech recognition, natural language processing, audio recognition, social network filtering, machine translation, bioinformatics, drug design and other fields. How to use deep learning methods, especially deep convolutional neural networks to solve the complex problems encountered in the tracking process, has great research value and research space.

This paper uses the powerful representation ability of convolutional neural network to improve the feature representation of video target tracking algorithm, appearance model modeling, motion model modeling, model update, etc., and effectively improve the performance of the algorithm. The main work and contributions are:

\begin{itemize}
\item{\textbf{ proposed a semantic information guided video target tracking algorithm. }We use the convolutional neural network to obtain the target \textbf{semantic information} to constrain the training process of the tracker, thereby improving the tracking effect. Specifically, first: We propose an example-guided correlation filter. Use the neural network to learn the instance-level semantic segmentation template of the image to constrain the learning of related filters. Secondly, in view of the complementarity between the semantic segmentation results of offline training and the correlation filtering results of online learning, we propose a self-correction mechanism for tracking results, using the segmentation results to correct the correlation filtering results. We have verified the effectiveness of these innovations in video target tracking applications on multiple challenging video target tracking databases. }
\item{\textbf{ proposed a video target tracking algorithm with enhanced spatial information. }Different from previous work, where the spatial search is done locally to determine the target location, we have introduced richer \textbf{spatial information}. Specifically, the tracking algorithm is optimized in terms of the motion model, and always perceives the position information of the object in the entire image plane, making up for the shortcomings of the limited target search range in the traditional local search mechanism, thereby effectively reducing the accumulated error and improving the robustness. Awesome. In order to further reduce the interference of similar objects, we propose an end-to-end training trajectory prediction module that can use the historical trajectory information of the object and the apparent information of the current frame to predict the possibility of the object appearing in each spatial position of the current frame . We verified the effectiveness of these innovations on multiple video tracking standard evaluation libraries, and greatly improved the accuracy and robustness of the tracking algorithm. }
\item{\textbf{ proposed a video target tracking algorithm enhanced by time information. } The algorithm mainly optimizes the video target tracking algorithm based on the twin network in feature extraction. First, we start from the problem of insufficient robustness of the online video tracking algorithm based on the twin network, and introduce \textbf{time information} into online video target tracking. Through the aggregation of the target appearance information from adjacent frames, the target appearance feature is enriched, and the video target tracking algorithm based on the twin network is limited to extracting the target appearance from a single frame, and thus the lack of ability to express the target appearance , Thereby improving the tracking effect and achieving robust tracking. On the basis of end-to-end time aggregation, we introduced an adversarial dropout module, and through end-to-end training on large-scale data sets, the Siamese network tracker can be used in situations where the target's appearance is not good due to motion blur, etc. Has better performance, thereby further improving the robustness of tracking. We have conducted algorithm comparison experiments and component analysis experiments on the currently popular video target tracking evaluation library to verify the effectiveness of the algorithm improvement. }
\item{\textbf{Propose an adaptive information-enhanced video target tracking algorithm} This algorithm mainly optimizes the video target tracking algorithm based on the twin network in the aspect of model adaptation. We introduced \textbf{adaptive information}. Previous work: offline training. Purpose: Improve the adaptive ability of the algorithm to specific objects. First, by slightly disturbing the pixels of the template image, the tracking performance of the twin network tracker for specific targets is improved. The adaptive information is obtained by back-propagating the gradient of the template image, and can be easily added to the existing twin network tracker in a pluggable manner without modifying the parameters of the network model. In online tracking, only a few times of gradient propagation and template image pixel value update are performed in the first frame to realize the fact tracking of the target. We also verified the effectiveness of the algorithm on multiple video target tracking evaluation libraries, and achieved good results in accuracy and real-time performance. }
\item{\textbf{Applying adversarial information to a video target tracking algorithm based on the Siamese network} We introduced \textbf{adversarial information} and proposed an adversarial attack algorithm. The purpose: to prove the robustness of the existing tracker Sex is not good. Specifically, we generate video-independent general disturbances for the video target tracking algorithm based on the twin network, which makes the tracker behave incorrectly. We add a small perturbation to the template image and a small patch to the search image so that the tracker predicts the location of the patch instead of the location of the target. Through the offline large-scale video target tracking data set training adversarial disturbance information, so as to achieve effective attacks. We verified the effectiveness of the proposed adversarial information on multiple video target tracking standard evaluation libraries, and at the same time verified its transferability in different backbone networks and different tracking frameworks. }
\end{itemize}

\KEYWORDS{Correlation Filter, Visual Object Tracking, Deep Learning, Conlvolutional Neural Network}% 英文关键词
%---------------------------------------------------------------------------%

>>>>

视频目标跟踪是计算机视觉领域中最重要和最具挑战性的研究课题之一，在智能监控、自动驾驶等领域有着潜在广阔的应用前景。目标跟踪的核心是估计图像序列的每帧中目标的运动状态。目标跟踪是计算机视觉领域的中层部分，为目标的行为理解提供了基础，因此具有非常重要的理论研究价值。同时，它具有广泛的实际应用，包括视频监控，交通流量监控，视频压缩和人机交互等。
然而，目标跟踪领域存在很多理论和技术问题有待解决，如运动模糊、光照变化、非刚性目标的形变、视角的变化导致的目标旋转、遮挡等。近年来深度学习的突破为解决目标跟踪中的一系列问题带来了可能。
深度学习是基于人工神经网络的机器学习方法。在过去的十年中，深度学习技术得到了飞速发展，已成功应用于计算机视觉，语音识别，自然语言处理，音频识别，社交网络过滤，机器翻译，生物信息学，药物设计等领域。如何利用深度学习方法，尤其是深度卷积神经网络解决跟踪过程中遇到的复杂问题，具有较大的研究价值和研究空间。

@@

\keywords{视频目标跟踪，深度学习，卷积神经网络，孪生网络，相关滤波}% 中文关键词
%-
%-> 英文摘要
%-
\intobmk\chapter*{Abstract}% 显示在书签但不显示在目录

Visual object tracking is one of the most important and challenging research topics in the field of computer vision, and it has potential and broad application prospects in the fields of intelligent surveillance and autonomous driving. The core of target tracking is to estimate the motion state of the target in each frame of the image sequence. Target tracking is the middle part of the computer vision field, which provides a basis for the understanding of target behavior, so it has very important theoretical research value. At the same time, it has a wide range of practical applications, including video surveillance, traffic flow monitoring, video compression, and human-computer interaction.
However, there are many theoretical and technical problems to be solved in the field of target tracking, such as motion blur, illumination changes, deformation of non-rigid targets, target rotation and occlusion caused by changes in perspective. In recent years, the breakthrough of deep learning has brought the possibility to solve a series of problems in target tracking.
Deep learning is a machine learning method based on artificial neural networks. In the past ten years, deep learning technology has been rapidly developed and has been successfully applied in computer vision, speech recognition, natural language processing, audio recognition, social network filtering, machine translation, bioinformatics, drug design and other fields. How to use deep learning methods, especially deep convolutional neural networks to solve the complex problems encountered in the tracking process, has great research value and research space.

This paper uses the powerful representation ability of convolutional neural network to improve the feature representation of video target tracking algorithm, appearance model modeling, motion model modeling, model adaptation, etc., and effectively improve the performance of the algorithm. At the same time, this article applies the adversarial information to the video target tracking task to study the robustness of the video target tracking algorithm relative to the adversarial samples. The main work and contributions of this paper are summarized as follows:

\begin{itemize}
\item{\textbf{We propose a video target tracking algorithm guided by semantic information.} The proposed algorithm uses the convolutional neural network to obtain the \textbf{semantic information} of the target, which is used to constrain the training process of the tracker, thereby improving the tracking effect. Specifically: First, we propose an instance-guided correlation filter, which uses a convolutional neural network to learn the semantic segmentation template at the instance level of the image, thereby constraining the learning of the correlation filter. Secondly, in view of the complementarity between the semantic segmentation results of offline training and the correlation filtering results of online learning, we propose a self-correction mechanism for tracking results, using the segmentation results to correct the correlation filtering results. We have verified the effectiveness of these innovations in video target tracking applications on multiple challenging video target tracking databases.}

\item{\textbf{We propose a visual object tracking algorithm with enhanced spatial information.} The proposed algorithm mainly optimizes the video target tracking algorithm based on the twin network in the aspect of motion model. Specifically: We have introduced richer \textbf{spatial information} for the twin tracking network, that is, always perceive the position information of the object in the entire image plane, which can make up for the shortcomings of the limited target search range in the traditional local search mechanism. Effectively reduce accumulated errors and improve robustness. In order to further reduce the interference of similar objects, we propose an end-to-end training trajectory prediction module that can use the historical trajectory information of the object and the apparent information of the current frame to predict the possibility of the target appearing in each spatial position of the current frame Sex. We have verified the effectiveness of these innovations on multiple video tracking standard evaluation libraries, and have greatly improved the accuracy and robustness of the tracking algorithm.}

\item{\textbf{We propose a visual object tracking algorithm enhanced by time information.} The proposed algorithm mainly optimizes the video target tracking algorithm based on twin network in feature extraction. First, we start from the problem of insufficient robustness of the online video tracking algorithm based on the twin network, and introduce \textbf{time information} into online video target tracking. Through the aggregation of the target appearance information from adjacent frames, the target appearance feature is enriched, and the video target tracking algorithm based on the twin network is limited to extracting the target appearance from a single frame, and thus the lack of ability to express the target appearance , Thereby improving the tracking effect and achieving robust tracking. On the basis of end-to-end time aggregation, we introduced the adversarial Dropout module, and through end-to-end training on large-scale data sets, so that the twin network tracker is not good in the appearance of the target due to motion blur, etc. Has better performance, thereby further improving the robustness of tracking. We have conducted algorithm comparison experiments and component analysis experiments on the currently popular video target tracking evaluation library to verify the effectiveness of the algorithm improvement.}

\item{\textbf{We propose an adaptive information-enhanced video target tracking algorithm} The proposed algorithm mainly optimizes the video target tracking algorithm based on the twin network in the aspect of model adaptation. We introduced \textbf{adaptive information} to the twin tracking network, which improves the tracking performance of the twin network tracker for specific targets by slightly perturbing the pixels of the template image. The adaptive information is obtained by performing gradient back propagation calculation on the template image, and can be easily added to the existing twin network tracker in a plug-and-play manner, without modifying the parameters of the network model. In online tracking, only a few times of gradient propagation and template image pixel value update in the first frame can achieve real-time target tracking. We also verified the effectiveness of the algorithm on multiple video target tracking evaluation libraries, and achieved good results in accuracy and real-time performance.}

\item{\textbf{We apply the adversarial information to the video target tracking algorithm based on the Siamese network.} The proposed algorithm applies \textbf{adversarial information} to the twin tracking network to study the robustness of the video target tracking algorithm relative to adversarial samples. Specifically, we generate video-independent general disturbances for the video target tracking algorithm based on the twin network, which makes the tracker behave incorrectly. The proposed adversarial disturbance information is obtained through offline large-scale video target tracking data set training, which can effectively attack any video while occupying very few computing resources. We verified the effectiveness of the proposed adversarial information on multiple video target tracking standard evaluation libraries, and at the same time verified its transferability in different backbone networks and different tracking frameworks.}
\end{itemize}

\KEYWORDS{Visual Object Tracking, Deep Learning, Conlvolutional Neural Network, Siamese Network, Correlation Filter}% 英文关键词
%---------------------------------------------------------------------------%
